Online (Bandit) Policy Selection for EASY-Backfilling

### 2017-04-03 paper/src/paper.tex
fix paper

The choice of a simulator is a critical part of experimental validation that
raises a tradeoff between precision and runtime. High precision can be obtained
by carefully modeling the platform and its network topology, or extracting
information from the jobs from their sources or post-morted logs. This is the
approach used by high-fidelity batch scheduling simulators such as
Batsim~\cite{batsim}. In the present work, the focus is on studying the
EASY-Backfilling mechanism in itself, without adressing the allocation problem.
Moreover, the experimental protocol used here requires many simulation runs.
Therefore, we set the precision/runtime tradeoff at the point which minimizes
simulation runtimes. We discard all topological information relative to the
platform and use the job processing times of the original workloads. In this
setup, the processors are considered to be undistinguishable from each other,
and jobs can be discontinuously mapped to any available processor on the
system. We develop a lightweight simulator\cite{ocst} and an accompanying
multi-armed bandit library\cite{obandit}. See the reproducibility paragraph
below for more information.  With this simulator, we are able to replay
EASY-FCFS on the CEA-CURIE trace in 33 seconds with a machine equipped with an
Intel(R) Core(TM) i5-4310U CPU @ 2.00GHz. As a point of comparison, the Batsim
simulator with no network communication modeling and no resource contiguity
takes more than 7 hours to replay the same trace on the same hardware. The
complete simulation campaign presented below is executed on a single Dell
PowerEdge R730 in 22 hours, including input preparation and analysis code.

@@

We use devise the following experimental campaign. For each of the 7 workloads
considered, we use the original data to resample 100 traces, each of a length
of two years. For each resampled trace, we run simulations using the
EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
threshold $\Theta$ at 40 hours, a sane default. The choice of this default
stems from the analysis developped in ~\cite{jsspp17}. This value is high
enough for queue reorderings to be leveraged while being low enough to provide
the functional requirement of preventing starvation. We simulate
EASY($P,\Theta=40h$) using the following strategies:

@@

The experiment campaign consists of 7 traces $\times$ 100 resamples $\times$
(12+2+2+2) algorithms, which equate to 12600 scheduling simulations spanning
two years.  This means a total simulated timespan of 25200 years. Moreover,
both the full-feedback and noisy feedback strategies will also be calling
simulations as part of their procedure. Indeed, these strategies re-simulate
the system 12 times every at every period $\Delta$, which bumps the total
effective simulation timespan to 92400 years. This motivates the use of a
lightweight simulator in the context of this research.

@@

We begin by attracting the attention of the reader to Table~\ref{tab:full},
which reports the average total waiting time performance improvement over
EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. More
precisely, let us denote the resampled traces of a given workload by $T_k$,
where k is the index of the resampled trace ($1 \le k \le 100$). We also denote
the number of jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by
$\bar{w}(T_k,S)$ the cumulative waiting time obtained by a simulation run for a
strategy $S$:

@@

\begin{equation}
  \frac{\sum_{k=1}^{k=100}\bar{w}(T_k,\text{EASY}(P_i,40h))-
  \bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}{\sum_{k=1}^{k=100}\bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}
\end{equation}

@@

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/full-ANL-Intr.pdf}
  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
    UniLu-Gaia trace. The average is obtained by resampling the original trace
    100 times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling.}
  \label{fig:all}
\end{figure}

@@

  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the FullFeedback, NoisyFeedback and EpsilonGreedy
    policies. The average is obtained by resampling the original trace 100
    times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling. Each figure is a different trace, and this
figure is followed-up in figure~\ref{fig:follow} for the UniLu-Gaia log.}

@@

\begin{itemize}
  \item Random $P \in P_i, 1 \le k \le 100$. ($\Delta=$1 week)
  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 week)
  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 day)
  \item Full (simulated) Feedback ($\Delta=$1 week)
  \item Noisy (simulated) Feedback ($\Delta=$1 week)
\end{itemize}

@@

  \caption{Share of the policies chosen by Epsilon-Greedy as a function of
  time. The average choice is obtained by resampling the original trace 100
times and aggregating by date.}

@@

  \caption{Share of the policies chosen by the Noisy policy. The average choice
  is obtained by resampling the original trace 100 times and aggregating by
  date.}

@@

All the code associated with this work is released under the ~\cite{isc}
license as a persistent zenodo~\cite{zenodo} archive at~\cite{zenodomain}.

@@

\textbf{one-liner : replication}

Running the experiments can be done on any platform equipped with the nix
package manager by running:

\begin{lstlisting}
nix-build (fetchTarball ...)
\end{lstlisting}

@@

All the software dependencies of our experiments are automatically managed by Nix.

@@

Nevertheless, the reader interested in using the tools outside the workflow can
obtain an environment by running the
\lstinline[basicstyle=\ttfamily]|nix-shell| command:

\begin{lstlisting}
nix-shell (fetchTarball ...)
\end{lstlisting}

Or alternatively, if the user is at the root of the extracted archive:

@@

Most simple workflow modifications (using different traces, changing the sample
sizes) can be obtained by modifying the
\lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|. For modification

>>>>

The choice of a simulator is a critical part of experimental validation that
raises a tradeoff between precision and runtime. High precision can be obtained
by carefully modeling the platform and its network topology, or extracting
information from the jobs from their sources or post-morted logs. This is the
approach used by high-fidelity batch scheduling simulators such as
Batsim~\cite{batsim}. In the present work, the focus is on studying the
EASY-Backfilling mechanism in itself, without adressing the allocation problem.
Moreover, the experimental protocol used here requires many simulation runs.
Therefore, we set the precision/runtime tradeoff at the point which minimizes
simulation runtimes. We discard all topological information relative to the
platform and use the job processing times of the original workloads. In this
setup, the processors are considered to be undistinguishable from each other,
and jobs can be discontinuously mapped to any available processor on the
system. We develop a lightweight simulator\cite{ocst} and release it to the
community. See the reproducibility paragraph below for more information.  With
this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in 33
seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
2.00GHz. As a point of comparison, the Batsim simulator with no network
communication modeling and no resource contiguity takes more than 7 hours to
replay the same trace on the same hardware. The complete simulation campaign
presented below is executed on a single Dell PowerEdge R730 in 22 hours,
including input preparation and analysis code.

@@

We use devise the following experimental campaign. For each of the 7 workloads
considered, we use the original data to resample 60 traces, each of a length
of two years. For each resampled trace, we run simulations using the
EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
threshold $\Theta$ at 40 hours, a sane default. The choice of this default
stems from the analysis developped in ~\cite{jsspp17}. This value is high
enough for queue reorderings to be leveraged while being low enough to provide
the functional requirement of preventing starvation. We simulate
EASY($P,\Theta=40h$) using the following strategies:

@@

The experiment campaign consists of 7 traces $\times$ 60 resamples $\times$
(12+2+2+2) algorithms, which equate to 12600 scheduling simulations spanning
two years.  This means a total simulated timespan of 25200 years. Moreover,
both the full-feedback and noisy feedback strategies will also be calling
simulations as part of their procedure. Indeed, these strategies re-simulate
the system 12 times every at every period $\Delta$, which bumps the total
effective simulation timespan to 92400 years. This motivates the use of a
lightweight simulator in the context of this research.

@@

We begin by attracting the attention of the reader to Table~\ref{tab:full},
which reports the average total waiting time performance improvement over
EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. More
precisely, let us denote the resampled traces of a given workload by $T_k$,
where k is the index of the resampled trace ($1 \le k \le 60$). We also denote
the number of jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by
$\bar{w}(T_k,S)$ the cumulative waiting time obtained by a simulation run for a
strategy $S$:

@@

\begin{equation}
  \frac{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(P_i,40h))-
  \bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}
\end{equation}

@@

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/full-ANL-Intr.pdf}
  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
    UniLu-Gaia trace. The average is obtained by resampling the original trace
    60 times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling.}
  \label{fig:all}
\end{figure}

@@

  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the FullFeedback, NoisyFeedback and EpsilonGreedy
    policies. The average is obtained by resampling the original trace 60 
    times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling. Each figure is a different trace, and this
figure is followed-up in figure~\ref{fig:follow} for the UniLu-Gaia log.}

@@

\begin{itemize}
  \item Random $P \in P_i, 1 \le k \le 60$. ($\Delta=$1 week)
  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 week)
  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 day)
  \item Full (simulated) Feedback ($\Delta=$1 week)
  \item Noisy (simulated) Feedback ($\Delta=$1 week)
\end{itemize}

@@

  \caption{Share of the policies chosen by Epsilon-Greedy as a function of
  time. The average choice is obtained by resampling the original trace 60 
times and aggregating by date.}

@@

  \caption{Share of the policies chosen by the Noisy policy. The average choice
  is obtained by resampling the original trace 60 times and aggregating by
  date.}

@@

All the code associated with this work is released under the ~\cite{isc}
license as a persistent zenodo~\cite{zenodo} archive~\cite{repro}.

@@

The archive containing the experimental workflow can be obtained at
\url{(omitted due to double-blind review)} Running the experiments can be done
on any platform equipped with the nix package manager by running at the root of
the extracted archive:

@@

All the software dependencies of our experiments are automatically managed by Nix.
The Nix software is available at \url{https://nixos.org/nix}.

@@

Nevertheless, the reader interested in using the tools outside the workflow can
obtain an environment by running a \lstinline[basicstyle=\ttfamily]|nix-shell| command
at the root of the extracted archive:

@@

Most simple workflow modifications (using different traces, changing the sample
sizes) can be obtained by modifying the
\lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile|.
